Contextuality and the Born rule in fuzzy quantum theories

### 2017-04-20 prop_letter.tex
prop_letter.tex \paragraph{Expectation Values for Quantum IVPMs} Updating...

Since our measurement is fuzzy, generally there are more than one
state consistent with a particular quantum IVPM. For example, consider
three intervals $\left[0,0\right]$, $\left[1,1\right]$ and \emph{$\left[0,1\right]$},
where $\left[0,0\right]$ and $\left[1,1\right]$ are called \imposs~and
\necess~as before, and \emph{$\left[0,1\right]$} is called \unknown~because
it provides no information. Then,
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\necess & \text{, if }P=\mathbb{1}\text{ ;}\\
\imposs & \text{, if }P=\mathbb{0}\text{ ;}\\
\unknown & \text{, otherwise.}
\end{cases}\label{eq:homomorphism-1}
\end{equation}
is a quantum IVPM. This measure represents the beliefs of an experimenter
with no prior knowledge about the particular quantum system in question,
and any state is consistent with $\bar{\mu}_{1}$ on any projectors.

@@

While mapping to $\mathscr{I}_{\infty}$ is one extreme case of quantum
IVPM, mapping to $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $
is another one, and corresponding to the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}.
However, before dive into the Kochen-Specker theorem, it is better
to understand how to define expectation value of an observable with
quantum IVPMs.

@@

Consider a quantum IVPM~$\bar{\mu}$ such that $l:\events\rightarrow\left[0,1\right]$
and $r:\events\rightarrow\left[0,1\right]$ are the left-end and the
right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[l\left(E\right),r\left(E\right)\right]$.
This time, when we express the observable~$\mathbf{O}$ as its spectrum
decomposition~$\sum_{i=1}^{N^{-}}\lambda_{i}^{-}P_{i}^{-}+\sum_{i=1}^{N^{+}}\lambda_{i}^{+}P_{i}^{+}$,
we separate the positive and negative eigenvalues and order them from
the smallest to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}\le\ldots\le\lambda_{1}^{-}\le0\le\lambda_{1}^{+}\le\ldots\le\lambda_{N^{+}}^{+}\textrm{ .}
\end{equation}
Let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$ and $\Delta f\left(P_{i}^{s\prime}\right)=f\left(P_{i}^{s\prime}\right)-f\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $f\in\left\{ l,r\right\} $.
We can combine the classical Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with (\ref{eq:quantum-expectation}) to define the expectation value
of $\mathbf{O}$ with respect to $\bar{\mu}$ as
\begin{eqnarray}
\expval{\mathbf{O}}_{\bar{\mu}} & = & \sum_{i=1}^{N^{+}}\lambda_{i}^{+}\left[\Delta l\left(P_{i}^{+\prime}\right),\Delta r\left(P_{i}^{+\prime}\right)\right]\\
 &  & +\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\left[\Delta l\left(P_{i}^{-\prime}\right),\Delta r\left(P_{i}^{-\prime}\right)\right]\textrm{ .}\label{eq:quantum-interval-expectation}
\end{eqnarray}
Let $\events'$ denote the set containing $\left\{ P_{i}^{-}\right\} _{i=1}^{N^{-}}\cup\left\{ P_{i}^{+}\right\} _{i=1}^{N^{+}}$
and satisfies the following conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$. 
\item $\mathbb{1}\in\events'$. 
\item For any projection~$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For a set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$. 
\end{itemize}
Then, we have the following theorem.

\begin{thm}Given a Hilbert space of $d\ge3$, for every quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
and any observable~$\mathbf{O}$, if $\bar{\mu}$ is convex, then
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\textrm{ .}\label{eq:quantum-interval-expectation-core}
\end{equation}
Conversely, if equation~(\ref{eq:quantum-interval-expectation-core})
holds for every observable, then $\bar{\mu}$ is convex.\end{thm}

>>>>

Since our measurement is fuzzy, generally there are more than one
state consistent with a particular quantum IVPM. For example, consider
three intervals $\left[0,0\right]$, $\left[1,1\right]$ and \emph{$\left[0,1\right]$},
where $\left[0,0\right]$ and $\left[1,1\right]$ are called \imposs~and
\necess~as before, and \emph{$\left[0,1\right]$} is called \unknown~because
it provides no information. Then,
\begin{equation}
\bar{\mu}_{1}\left(P\right)=\begin{cases}
\necess & \text{, if }P=\mathbb{1}\text{ ;}\\
\imposs & \text{, if }P=\mathbb{0}\text{ ;}\\
\unknown & \text{, otherwise,}
\end{cases}\label{eq:homomorphism-1}
\end{equation}
is a quantum IVPM. This measure represents the beliefs of an experimenter
with no prior knowledge about the particular quantum system in question,
and any state is consistent with $\bar{\mu}_{1}$ on any projectors.

@@

While mapping to $\mathscr{I}_{\infty}$ is one extreme case of quantum
IVPM, mapping to $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $
is another one, and corresponding to the Kochen-Specker theorem~\cite{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}.
However, before diving into the Kochen-Specker theorem, it is better
to understand how to define expectation value of an observable with
quantum IVPMs.

@@

Consider a quantum IVPM~$\bar{\mu}$ such that $\boldsymbol{l}:\events\rightarrow\left[0,1\right]$
and $\boldsymbol{r}:\events\rightarrow\left[0,1\right]$ are the left-end
and the right-end of $\bar{\mu}$, respectively, i.e., $\bar{\mu}\left(E\right)=\left[\boldsymbol{l}\left(E\right),\boldsymbol{r}\left(E\right)\right]$.
This time, when we express the observable~$\mathbf{O}$ as its spectrum
decomposition~$\sum_{i=1}^{N^{-}}\lambda_{i}^{-}P_{i}^{-}+\sum_{i=1}^{N^{+}}\lambda_{i}^{+}P_{i}^{+}$,
we separate the positive and negative eigenvalues and order them from
the smallest to the largest 
\begin{equation}
\lambda_{N^{-}}^{-}\le\ldots\le\lambda_{1}^{-}\le0\le\lambda_{1}^{+}\le\ldots\le\lambda_{N^{+}}^{+}\textrm{ .}
\end{equation}
Let $P_{i}^{s\prime}=\sum_{j\ge i}P_{j}^{s}$ and $\Delta f\left(P_{i}^{s\prime}\right)=f\left(P_{i}^{s\prime}\right)-f\left(P_{i+1}^{s\prime}\right)$,
where $s\in\left\{ +,-\right\} $ and $f\in\left\{ \boldsymbol{l},\boldsymbol{r}\right\} $.
We can combine the classical Choquet integral~\cite{Vitali1925,Choquet1954,GilboaSchmeidler1994,Grabisch2016}
with Eq.~(\ref{eq:quantum-expectation}) to define the expectation
value of $\mathbf{O}$ with respect to $\bar{\mu}$ as
\begin{eqnarray}
\expval{\mathbf{O}}_{\bar{\mu}} & = & \sum_{i=1}^{N^{+}}\lambda_{i}^{+}\left[\Delta\boldsymbol{l}\left(P_{i}^{+\prime}\right),\Delta\boldsymbol{r}\left(P_{i}^{+\prime}\right)\right]\nonumber \\
 &  & +\sum_{i=1}^{N^{-}}\lambda_{i}^{-}\left[\Delta\boldsymbol{l}\left(P_{i}^{-\prime}\right),\Delta\boldsymbol{r}\left(P_{i}^{-\prime}\right)\right]\textrm{ .}\label{eq:quantum-interval-expectation}
\end{eqnarray}
Let $\events'$ denote the set containing $\left\{ P_{i}^{-}\right\} _{i=1}^{N^{-}}\cup\left\{ P_{i}^{+}\right\} _{i=1}^{N^{+}}$
and satisfies the following conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$. 
\item $\mathbb{1}\in\events'$. 
\item For any projection~$P\in\events'$, we have $\mathbb{1}-P\in\events'$. 
\item For a set of mutually orthogonal projections~$\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$. 
\end{itemize}
And $\bar{\mu}$ is called convex if for all commuting $P_{0},P_{1}\in\events$,
$\bar{\mu}\left(P_{0}\vee P_{1}\right)+\bar{\mu}\left(P_{0}P_{1}\right)\subseteq\bar{\mu}\left(P_{0}\right)+\bar{\mu}\left(P_{1}\right)$,
where $P_{0}\vee P_{1}=P_{0}+P_{1}-P_{0}P_{1}$~\cite{Griffiths2003}.
Then, we have the following theorem.

\begin{thm}Given a Hilbert space of $d\ge3$, for every quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$
and any observable~$\mathbf{O}$, if $\bar{\mu}$ is convex, then
\begin{equation}
\expval{\mathbf{O}}_{\bar{\mu}}=\left[\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho},\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho}\right]\textrm{ .}\label{eq:quantum-interval-expectation-core}
\end{equation}
Conversely, if Eq.~(\ref{eq:quantum-interval-expectation-core})
holds for every observable, then $\bar{\mu}$ is convex.\end{thm}

For example, to compute the expectation value of $\sigma_{\mathrm{z}}=\proj{0}-\proj{1}$
with respect to $\bar{\mu}_{1}$, we have $N^{+}=N^{-}=1$, $\lambda_{1}^{+}=1$,
$\lambda_{1}^{-}=-1$, $P_{1}^{+\prime}=\proj{0}$, $P_{1}^{-\prime}=\proj{1}$,
$P_{2}^{+\prime}=P_{2}^{-\prime}=\mathbb{0}$. Let $\left[\boldsymbol{l}_{1}\left(E\right),\boldsymbol{r}_{1}\left(E\right)\right]=\bar{\mu}_{1}\left(E\right)$,
then $\Delta\boldsymbol{l}_{1}\left(P_{1}^{+\prime}\right)=\boldsymbol{l}_{1}\left(\proj{0}\right)-\boldsymbol{l}_{1}\left(\mathbb{0}\right)=0$,
$\Delta\boldsymbol{l}_{1}\left(P_{1}^{-\prime}\right)=\boldsymbol{l}_{1}\left(\proj{1}\right)-\boldsymbol{l}_{1}\left(\mathbb{0}\right)=0$,
$\Delta\boldsymbol{r}_{1}\left(P_{1}^{+\prime}\right)=\boldsymbol{r}_{1}\left(\proj{0}\right)-\boldsymbol{r}_{1}\left(\mathbb{0}\right)=1$,
$\Delta\boldsymbol{r}_{1}\left(P_{1}^{-\prime}\right)=\boldsymbol{r}_{1}\left(\proj{1}\right)-\boldsymbol{r}_{1}\left(\mathbb{0}\right)=1$.
Therefore, $\expval{\sigma_{\mathrm{z}}}_{\bar{\mu}_{1}}=\left[0,1\right]+\left(-1\right)\left[0,1\right]=\left[-1,1\right]$.

On the other hand, it is easy to verify that $\bar{\mu}_{1}$ is convex.
Every state is consistent with $\bar{\mu}_{1}$, in particular $\proj{0}$
and $\proj{1}$. It is easy to see 
\begin{eqnarray*}
\min_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho} & = & \expval{\sigma_{\mathrm{z}}}_{\proj{1}}=-1\textrm{ ,}\\
\max_{\rho\in\coreBorn\left(\bar{\mu},\events'\right)}\expval{\mathbf{O}}_{\rho} & = & \expval{\sigma_{\mathrm{z}}}_{\proj{0}}=1\textrm{ ,}
\end{eqnarray*}
which is consistent with the theorem.
